---
title: "Analysis of Households with Negative Net Worth Using Data from the 2019 Survey of Consumer Finances"
author: Shehryar Nabi, Senior Research Associate, Aspen Institute Financial Security Program
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Set working directory
setwd(paste0(getwd()))


# Load libraries
library(tidyverse)
library(ggplot2)
library(spatstat)
library(survey)
library(magrittr)
library(expss)
library(weights)

# Read in data

# 1989

scf_89_master <- read.csv("SCFP1989.csv")
scf_89 <- scf_89_master

# Debt

otherdebt <- scf_89 %>%
  select(OTHLOC, OTH_INST, ODEBT)
scf_89$OTHDBT <- rowSums(otherdebt)

# Assets 

securities <- select(scf_89, c(NMMF, SAVBND, STOCKS, BOND, WGT))
other_fin <- select(scf_89, c(CDS, CASHLI, OTHMA, OTHFIN, WGT))
nonres <- select(scf_89, c(ORESRE, NNRESRE, WGT))

scf_89$SECURITIES <- rowSums(securities[1:4])
scf_89$OTHFIN_COMPLETE <- rowSums(other_fin[1:4])
scf_89$NONRES <- rowSums(nonres[1:2])


# 1992

scf_92_master <- read.csv("SCFP1992.csv")
scf_92 <- scf_92_master

# Debt

otherdebt <- scf_92 %>%
  select(OTHLOC, OTH_INST, ODEBT)
scf_92$OTHDBT <- rowSums(otherdebt)

# Assets 

securities <- select(scf_92, c(NMMF, SAVBND, STOCKS, BOND, WGT))
other_fin <- select(scf_92, c(CDS, CASHLI, OTHMA, OTHFIN, WGT))
nonres <- select(scf_92, c(ORESRE, NNRESRE, WGT))

scf_92$SECURITIES <- rowSums(securities[1:4])
scf_92$OTHFIN_COMPLETE <- rowSums(other_fin[1:4])
scf_92$NONRES <- rowSums(nonres[1:2])


# 1995

scf_95_master <- read.csv("SCFP1995.csv")
scf_95 <- scf_95_master

# Debt

otherdebt <- scf_95 %>%
  select(OTHLOC, OTH_INST, ODEBT)
scf_95$OTHDBT <- rowSums(otherdebt)

# Assets 

securities <- select(scf_95, c(NMMF, SAVBND, STOCKS, BOND, WGT))
other_fin <- select(scf_95, c(CDS, CASHLI, OTHMA, OTHFIN, WGT))
nonres <- select(scf_95, c(ORESRE, NNRESRE, WGT))

scf_95$SECURITIES <- rowSums(securities[1:4])
scf_95$OTHFIN_COMPLETE <- rowSums(other_fin[1:4])
scf_95$NONRES <- rowSums(nonres[1:2])


# 1998

scf_98_master <- read.csv("SCFP1998.csv")
scf_98 <- scf_98_master

# Debt

otherdebt <- scf_98 %>%
  select(OTHLOC, OTH_INST, ODEBT)
scf_98$OTHDBT <- rowSums(otherdebt)

# Assets 

securities <- select(scf_98, c(NMMF, SAVBND, STOCKS, BOND, WGT))
other_fin <- select(scf_98, c(CDS, CASHLI, OTHMA, OTHFIN, WGT))
nonres <- select(scf_98, c(ORESRE, NNRESRE, WGT))

scf_98$SECURITIES <- rowSums(securities[1:4])
scf_98$OTHFIN_COMPLETE <- rowSums(other_fin[1:4])
scf_98$NONRES <- rowSums(nonres[1:2])


# 2001

scf_01_master <- read.csv("SCFP2001.csv")
scf_01 <- scf_01_master

# Debt

otherdebt <- scf_01 %>%
  select(OTHLOC, OTH_INST, ODEBT)
scf_01$OTHDBT <- rowSums(otherdebt)

# Assets 

securities <- select(scf_01, c(NMMF, SAVBND, STOCKS, BOND, WGT))
other_fin <- select(scf_01, c(CDS, CASHLI, OTHMA, OTHFIN, WGT))
nonres <- select(scf_01, c(ORESRE, NNRESRE, WGT))

scf_01$SECURITIES <- rowSums(securities[1:4])
scf_01$OTHFIN_COMPLETE <- rowSums(other_fin[1:4])
scf_01$NONRES <- rowSums(nonres[1:2])


# 2004

scf_04_master <- read.csv("SCFP2004.csv")
scf_04 <- scf_04_master

# Debt

otherdebt <- scf_04 %>%
  select(OTHLOC, OTH_INST, ODEBT)
scf_04$OTHDBT <- rowSums(otherdebt)

# Assets 

securities <- select(scf_04, c(NMMF, SAVBND, STOCKS, BOND, WGT))
other_fin <- select(scf_04, c(CDS, CASHLI, OTHMA, OTHFIN, WGT))
nonres <- select(scf_04, c(ORESRE, NNRESRE, WGT))

scf_04$SECURITIES <- rowSums(securities[1:4])
scf_04$OTHFIN_COMPLETE <- rowSums(other_fin[1:4])
scf_04$NONRES <- rowSums(nonres[1:2])


# 2007
 
scf_07_master <- read.csv("SCFP2007.csv")
scf_07 <- scf_07_master

# Debt

otherdebt <- scf_07 %>%
  select(OTHLOC, OTH_INST, ODEBT)
scf_07$OTHDBT <- rowSums(otherdebt)

# Assets 

securities <- select(scf_07, c(NMMF, SAVBND, STOCKS, BOND, WGT))
other_fin <- select(scf_07, c(CDS, CASHLI, OTHMA, OTHFIN, WGT))
nonres <- select(scf_07, c(ORESRE, NNRESRE, WGT))

scf_07$SECURITIES <- rowSums(securities[1:4])
scf_07$OTHFIN_COMPLETE <- rowSums(other_fin[1:4])
scf_07$NONRES <- rowSums(nonres[1:2])


# 2010

scf_10_master <- read.csv("SCFP2010.csv")
scf_10 <- scf_10_master

# Debt

otherdebt <- scf_10 %>%
  select(OTHLOC, OTH_INST, ODEBT)
scf_10$OTHDBT <- rowSums(otherdebt)

# Assets 

securities <- select(scf_10, c(NMMF, SAVBND, STOCKS, BOND, WGT))
other_fin <- select(scf_10, c(CDS, CASHLI, OTHMA, OTHFIN, WGT))
nonres <- select(scf_10, c(ORESRE, NNRESRE, WGT))

scf_10$SECURITIES <- rowSums(securities[1:4])
scf_10$OTHFIN_COMPLETE <- rowSums(other_fin[1:4])
scf_10$NONRES <- rowSums(nonres[1:2])


# 2013

scf_13_master <- read.csv("SCFP2013.csv")
scf_13 <- scf_13_master

# Debt

otherdebt <- scf_13 %>%
  select(OTHLOC, OTH_INST, ODEBT)
scf_13$OTHDBT <- rowSums(otherdebt)

# Assets 

securities <- select(scf_13, c(NMMF, SAVBND, STOCKS, BOND, WGT))
other_fin <- select(scf_13, c(CDS, CASHLI, OTHMA, OTHFIN, WGT))
nonres <- select(scf_13, c(ORESRE, NNRESRE, WGT))

scf_13$SECURITIES <- rowSums(securities[1:4])
scf_13$OTHFIN_COMPLETE <- rowSums(other_fin[1:4])
scf_13$NONRES <- rowSums(nonres[1:2])


# 2016

scf_16_master <- read.csv("SCFP2016.csv")
scf_16 <- scf_16_master

# Debt

otherdebt <- scf_16 %>%
  select(OTHLOC, OTH_INST, ODEBT)
scf_16$OTHDBT <- rowSums(otherdebt)

# Assets 

securities <- select(scf_16, c(NMMF, SAVBND, STOCKS, BOND, WGT))
other_fin <- select(scf_16, c(CDS, CASHLI, OTHMA, OTHFIN, WGT))
nonres <- select(scf_16, c(ORESRE, NNRESRE, WGT))

scf_16$SECURITIES <- rowSums(securities[1:4])
scf_16$OTHFIN_COMPLETE <- rowSums(other_fin[1:4])
scf_16$NONRES <- rowSums(nonres[1:2])


# 2019

scf_19_master <- read.csv("SCFP2019.csv")
scf_19 <- scf_19_master

# Debt

otherdebt <- scf_19 %>%
  select(OTHLOC, OTH_INST, ODEBT)
scf_19$OTHDBT <- rowSums(otherdebt)

# Assets 

securities <- select(scf_19, c(NMMF, SAVBND, STOCKS, BOND, WGT))
other_fin <- select(scf_19, c(CDS, CASHLI, OTHMA, OTHFIN, WGT))
nonres <- select(scf_19, c(ORESRE, NNRESRE, WGT))

scf_19$SECURITIES <- rowSums(securities[1:4])
scf_19$OTHFIN_COMPLETE <- rowSums(other_fin[1:4])
scf_19$NONRES <- rowSums(nonres[1:2])


# Disable scientific notation
options(scipen = 999)


# Functions # 

# Faster weighted quantile function
weighted.quant <- function(var, weight, quantile) {
  weighted.quantile(var, weight, probs = seq(0, 1, quantile))
}


# Get weighted net worth quantile breaks
weighted_networth_quantiles <- function(var, weight, quantile) {
  df <- data.frame(weighted.quant(var, weight, quantile))
  return(c(df[,1]))
}


# Rename race categories

scf_19$RACE[scf_19$RACE == 1] <- "White, non-Hispanic"
scf_19$RACE[scf_19$RACE == 2] <- "Black, non-Hispanic"
scf_19$RACE[scf_19$RACE == 3] <- "Hispanic"
scf_19$RACE[scf_19$RACE == 5] <- "Other"



# Create income quintiles 
inc_quintile_breaks_19 <- c(weighted_networth_quantiles(scf_19$INCOME, 
                                                  scf_19$WGT, 1/5))
inc_quintile_breaks_19[1] <- inc_quintile_breaks_19[1] - 1
scf_19$INCOME_quintiles <- cut(scf_19$INCOME,
                            c(inc_quintile_breaks_19),
                            labels = c("0-19.9", "20-39.9",
                                   "40-59.9", "60-79.9", 
                                   "80-100"))


# SCF net worth quintiles 
scf_19$nw_scf_breaks <- cut(scf_19$NETWORTH,
    c(-955502, 12410, 121760, 404100, 1218737, 1967199000),
    labels = c("Less than 25", "25-49.9",
                "50-74.9", "75-89.9", 
                "90-100"), na.rm = TRUE)


# Create age groups 

scf_19$age_group <- cut(scf_19$AGE,
                            c(17, 34, 54, 96),
                            labels = c("Under 35", "35-54",
                                   "Over 55"))

```

# Solutions Scan

## Share of households with negative net worth for each year from 1989-2019

```{r share net debtors of population}

# 1989 (below process repeated for each year)

# New binary variable that distinguishes net debt HHs from all HHs
scf_89 <- mutate(scf_89, 
                 netdebt_status = ifelse(scf_89$NETWORTH < 0, 1, 0))

# Taking weighted share of net debt households and all else
wpct(scf_89$netdebt_status, scf_89$WGT)


# 1992

scf_92 <- mutate(scf_92, 
                 netdebt_status = ifelse(scf_92$NETWORTH < 0, 1, 0))

wpct(scf_92$netdebt_status, scf_92$WGT)


# 1995

scf_95 <- mutate(scf_95, 
                 netdebt_status = ifelse(scf_95$NETWORTH < 0, 1, 0))

wpct(scf_95$netdebt_status, scf_95$WGT)


# 1998

scf_98 <- mutate(scf_98, 
                 netdebt_status = ifelse(scf_98$NETWORTH < 0, 1, 0))

wpct(scf_98$netdebt_status, scf_98$WGT)


# 2001

scf_01 <- mutate(scf_01, 
                 netdebt_status = ifelse(scf_01$NETWORTH < 0, 1, 0))

wpct(scf_01$netdebt_status, scf_01$WGT)


# 2004

scf_04 <- mutate(scf_04, 
                 netdebt_status = ifelse(scf_04$NETWORTH < 0, 1, 0))

wpct(scf_04$netdebt_status, scf_04$WGT)


# 2007

scf_07 <- mutate(scf_07, 
                 netdebt_status = ifelse(scf_07$NETWORTH < 0, 1, 0))

wpct(scf_07$netdebt_status, scf_07$WGT)


# 2010 

scf_10 <- mutate(scf_10, 
                 netdebt_status = ifelse(scf_10$NETWORTH < 0, 1, 0))

wpct(scf_10$netdebt_status, scf_10$WGT)


# 2013 

scf_13 <- mutate(scf_13, 
                 netdebt_status = ifelse(scf_13$NETWORTH < 0, 1, 0))

wpct(scf_13$netdebt_status, scf_13$WGT)


# 2016

scf_16 <- mutate(scf_16, 
                 netdebt_status = ifelse(scf_16$NETWORTH < 0, 1, 0))

wpct(scf_16$netdebt_status, scf_16$WGT)


# 2019

scf_19 <- mutate(scf_19, 
                 netdebt_status = ifelse(scf_19$NETWORTH < 0, 1, 0))

wpct(scf_19$netdebt_status, scf_19$WGT)
```


## Income and net worth

I calculate the differences in median income, assets, debt, and net worth between negative net worth households and all households. I also calculate the share of households with negative net worth by income quintile. 

```{r Income and net worth}

# Comparing median income of households with and without net debt

weighted.median(neg_nw_19$INCOME, neg_nw_19$WGT)
weighted.median(scf_19$INCOME, scf_19$WGT)


# Getting median debt, asset and net worth values for all HHs with net debt

neg_nw_19_wealth <- neg_nw_19 %>%
  group_by(INCOME_quintiles) %>%
  summarise(median_debt = weighted.median(DEBT, WGT)/1000,
            median_assets = weighted.median(ASSET, WGT)/1000,
            median_nw = weighted.median(NETWORTH, WGT)/1000, 
            n = n())


# Getting share of net debtors by income quintile

neg_nw_wealth$prop <- wpct(neg_nw_19$INCOME_quintiles, neg_nw_19$WGT)


# Exporting data

write.csv(neg_nw_wealth, paste(getwd(), '\\csv-exports', 
                '\\neg_nw_wealth.csv', 
                sep=''))

```


## Racial breakdown of households with negative net worth

```{r racial breakdown}

# Data frame comparing racial breakdown of all HHs vs net debtors

race_scf_netdebtors <- data.frame(scf = wpct(scf_19$RACE, 
                                             scf_19$WGT),
                                  net_debtors = wpct(neg_nw_19$RACE,
                                                     neg_nw_19$WGT))


# Exporting data 

write.csv(race_scf_netdebtors,  paste(getwd(), '\\csv-exports', 
                '\\race_scf_netdebtors.csv', 
                sep=''))

```


## Age and family structure of households with negative net worth compared to the overall population

See code descriptions for family structure here: https://sda.berkeley.edu/sdaweb/docs/scfcomb2019/DOC/hcbk0001.htm#FAMSTRUCT. 

```{r age fam}

# Age 

weighted.median(scf_19$AGE, scf_19$WGT)
weighted.median(neg_nw_19$AGE, neg_nw_19$WGT)


# Family structure

wpct(scf_19$FAMSTRUCT, scf_19$WGT)
wpct(neg_nw_19$FAMSTRUCT, neg_nw_19$WGT)


# Gender

wpct(scf_19$HHSEX, scf_19$WGT)
wpct(neg_nw_19$HHSEX, neg_nw_19$WGT)


# Gender breakdown of single parents

wpct(filter(scf_19, scf_19$FAMSTRUCT == 1)$HHSEX,
     filter(scf_19, scf_19$FAMSTRUCT == 1)$WGT)
wpct(filter(neg_nw_19, neg_nw_19$FAMSTRUCT == 1)$HHSEX,
     filter(neg_nw_19, neg_nw_19$FAMSTRUCT == 1)$WGT)


# Family structure breakdown of women-respondent HHs

wpct(filter(scf_19, scf_19$HHSEX == 2)$FAMSTRUCT,
     filter(scf_19, scf_19$HHSEX == 2)$WGT)
wpct(filter(neg_nw_19, neg_nw_19$HHSEX == 2)$FAMSTRUCT,
     filter(neg_nw_19, neg_nw_19$HHSEX == 2)$WGT)


```


## Debt holding rates 

I calculate the proportion of negative net worth households holding different types of debt. 

```{r neg nw debts}

debt_hold <- data.frame(RESDBT = wpct(neg_nw_19$RESDBT > 0,
                                      neg_nw_19$WGT)[1],
                        MRTHEL = wpct(neg_nw_19$MRTHEL > 0, 
                                      neg_nw_19$WGT)[1],
                        CCBAL = wpct(neg_nw_19$CCBAL > 0, 
                                     neg_nw_19$WGT)[1],
                        EDN_INST = wpct(neg_nw_19$EDN_INST > 0, 
                                        neg_nw_19$WGT)[1],
                        VEH_INST = wpct(neg_nw_19$VEH_INST > 0, 
                                        neg_nw_19$WGT)[1],
                        OTHDBT = wpct(neg_nw_19$OTHDBT > 0, 
                                      neg_nw_19$WGT)[1])


# Exporting data

write.csv(debt_hold, paste(getwd(), '\\csv-exports', 
                '\\debt_hold.csv', 
                sep=''))

```


## Median debt

I calculate the median value per debt type among households holding it for both households with negative net worth and overall.

```{r Debt mix among people holding debt}

# Calculating median debt value for select debt types among net debtors 
# holding them

debts_holding <- data.frame(RESDBT = weighted.median(filter(neg_nw_19,neg_nw_19$RESDBT > 0)$RESDBT, 
                filter(neg_nw_19,neg_nw_19$RESDBT > 0)$WGT),
                            MRTHEL = 
weighted.median(filter(neg_nw_19,neg_nw_19$MRTHEL > 0)$MRTHEL, 
                filter(neg_nw_19,neg_nw_19$MRTHEL > 0)$WGT),
                            CCBAL = 
weighted.median(filter(neg_nw_19,neg_nw_19$CCBAL > 0)$CCBAL, 
                filter(neg_nw_19,neg_nw_19$CCBAL > 0)$WGT),
                            EDN_INST = 
weighted.median(filter(neg_nw_19,neg_nw_19$EDN_INST > 0)$EDN_INST, 
                filter(neg_nw_19,neg_nw_19$EDN_INST > 0)$WGT),
                            VEH_INST = 
weighted.median(filter(neg_nw_19,neg_nw_19$VEH_INST > 0)$VEH_INST, 
                filter(neg_nw_19,neg_nw_19$VEH_INST > 0)$WGT),
                            OTHDBT = 
weighted.median(filter(neg_nw_19,neg_nw_19$OTHDBT > 0)$OTHDBT, 
                filter(neg_nw_19,neg_nw_19$OTHDBT > 0)$WGT))                              

# Export to csv 

write.csv(debts_holding,  paste(getwd(), '\\csv-exports', 
                '\\debts_holding.csv', 
                sep=''))

```


# Supplementary Figures

## Net debtors by age group and race

I find the age distribution of households with negative net worth compared to overall households. I also calculate the proportion of households with net debt by race and age. 

```{r Net debtors by age group and race}

# Getting the distribution of ages for negative net worth HHs and overall

wpct(scf_19$age_group, scf_19$WGT)
wpct(neg_nw_19$age_group, neg_nw_19$WGT)

# Finding share of HHs with net debt and debt types by age group and race

prop_nd_race <- scf_19 %>%
  group_by(age_group, RACE) %>%
  summarise(prop = wpct(NETWORTH < 0, WGT)[1],
            prop_edu_all = wpct(EDN_INST > 0, WGT)[1],
            prop_edu = wpct(NETWORTH < 0 & EDN_INST > 0, WGT)[1],
            prop_veh = wpct(NETWORTH < 0 & VEH_INST > 0, WGT)[1],
            prop_ccbal = wpct(NETWORTH < 0 & CCBAL > 0, WGT)[1],
            prop_mort = wpct(NETWORTH < 0 & MRTHEL > 0, WGT)[1],
            prop_rdebt = wpct(NETWORTH < 0 & RESDBT > 0, WGT)[1],
            count = n())


# Exporting to CSV 

write.csv(prop_nd_race, paste(getwd(), '\\csv-exports', 
                '\\prop_nd_race.csv', 
                sep=''))

```


## Median asset values for negative net worth households holding the asset in 2019

```{r Asset mix for people with net debt and holding asset}


assets_holding_19 <- data.frame(LIQ = weighted.median(filter(neg_nw_19, neg_nw_19$LIQ >0)$LIQ,filter(neg_nw_19, neg_nw_19$LIQ > 0)$WGT),
                            SECURITIES = 
weighted.median(filter(neg_nw_19,neg_nw_19$SECURITIES > 0)$SECURITIES, 
                filter(neg_nw_19,neg_nw_19$SECURITIES > 0)$WGT),
                            RETQLIQ = 
weighted.median(filter(neg_nw_19,neg_nw_19$RETQLIQ > 0)$RETQLIQ, 
                filter(neg_nw_19,neg_nw_19$RETQLIQ > 0)$WGT),
                            OTHFIN_COMPLETE = 
weighted.median(filter(neg_nw_19,neg_nw_19$OTHFIN_COMPLETE > 0)$OTHFIN_COMPLETE, 
                filter(neg_nw_19,neg_nw_19$OTHFIN_COMPLETE > 0)$WGT),
                            VEHIC = 
weighted.median(filter(neg_nw_19,neg_nw_19$VEHIC> 0)$VEHIC, 
                filter(neg_nw_19,neg_nw_19$VEHIC > 0)$WGT),
                            HOUSES = 
weighted.median(filter(neg_nw_19,neg_nw_19$HOUSES > 0)$HOUSES, 
                filter(neg_nw_19,neg_nw_19$HOUSES > 0)$WGT),
                            NONRES = 
weighted.median(filter(neg_nw_19,neg_nw_19$NONRES > 0)$NONRES, 
                filter(neg_nw_19,neg_nw_19$NONRES > 0)$WGT),
                            BUS = 
weighted.median(filter(neg_nw_19,neg_nw_19$BUS > 0)$BUS, 
                filter(neg_nw_19,neg_nw_19$BUS > 0)$WGT),
                            OTHNFIN = 
weighted.median(filter(neg_nw_19,neg_nw_19$OTHNFIN > 0)$OTHNFIN, 
                filter(neg_nw_19,neg_nw_19$OTHNFIN > 0)$WGT))  

```


## Median assets for negative net worth households overall by year

```{r median assets overall}

assets_med_19 <- data.frame(transaction = weighted.median(neg_nw_19$LIQ,
                                                          neg_nw_19$WGT)/1000,
                  securities = weighted.median(neg_nw_19$SECURITIES, 
                                               neg_nw_19$WGT)/1000,
                   ret = weighted.median(neg_nw_19$RETQLIQ, 
                                         neg_nw_19$WGT)/1000,
                   othfin = weighted.median(neg_nw_19$OTHFIN_COMPLETE,
                                            neg_nw_19$WGT)/1000,
                   vehic = weighted.median(neg_nw_19$VEHIC, 
                                           neg_nw_19$WGT)/1000,
                   res = weighted.median(neg_nw_19$HOUSES, 
                                         neg_nw_19$WGT)/1000,
                   nonres = weighted.median(neg_nw_19$NONRES, 
                                            neg_nw_19$WGT)/1000,
                   business = weighted.median(neg_nw_19$BUS, 
                                              neg_nw_19$WGT)/1000,
                   othnfin = weighted.median(neg_nw_19$OTHNFIN, 
                                             neg_nw_19$WGT)/1000)


assets_med_16 <- data.frame(transaction = weighted.median(neg_nw_16$LIQ,
                                                          neg_nw_16$WGT)/1000,
                  securities = weighted.median(neg_nw_16$SECURITIES, 
                                               neg_nw_16$WGT)/1000,
                   ret = weighted.median(neg_nw_16$RETQLIQ, 
                                         neg_nw_16$WGT)/1000,
                   othfin = weighted.median(neg_nw_16$OTHFIN_COMPLETE,
                                            neg_nw_16$WGT)/1000,
                   vehic = weighted.median(neg_nw_16$VEHIC, 
                                           neg_nw_16$WGT)/1000,
                   res = weighted.median(neg_nw_16$HOUSES, 
                                         neg_nw_16$WGT)/1000,
                   nonres = weighted.median(neg_nw_16$NONRES, 
                                            neg_nw_16$WGT)/1000,
                   business = weighted.median(neg_nw_16$BUS, 
                                              neg_nw_16$WGT)/1000,
                   othnfin = weighted.median(neg_nw_16$OTHNFIN, 
                                             neg_nw_16$WGT)/1000)


assets_med_13 <- data.frame(transaction = weighted.median(neg_nw_13$LIQ,
                                                          neg_nw_13$WGT)/1000,
                  securities = weighted.median(neg_nw_13$SECURITIES, 
                                               neg_nw_13$WGT)/1000,
                   ret = weighted.median(neg_nw_13$RETQLIQ, 
                                         neg_nw_13$WGT)/1000,
                   othfin = weighted.median(neg_nw_13$OTHFIN_COMPLETE,
                                            neg_nw_13$WGT)/1000,
                   vehic = weighted.median(neg_nw_13$VEHIC, 
                                           neg_nw_13$WGT)/1000,
                   res = weighted.median(neg_nw_13$HOUSES, 
                                         neg_nw_13$WGT)/1000,
                   nonres = weighted.median(neg_nw_13$NONRES, 
                                            neg_nw_13$WGT)/1000,
                   business = weighted.median(neg_nw_13$BUS, 
                                              neg_nw_13$WGT)/1000,
                   othnfin = weighted.median(neg_nw_13$OTHNFIN, 
                                             neg_nw_13$WGT)/1000)


assets_med_10 <- data.frame(transaction = weighted.median(neg_nw_10$LIQ,
                                                          neg_nw_10$WGT)/1000,
                  securities = weighted.median(neg_nw_10$SECURITIES, 
                                               neg_nw_10$WGT)/1000,
                   ret = weighted.median(neg_nw_10$RETQLIQ, 
                                         neg_nw_10$WGT)/1000,
                   othfin = weighted.median(neg_nw_10$OTHFIN_COMPLETE,
                                            neg_nw_10$WGT)/1000,
                   vehic = weighted.median(neg_nw_10$VEHIC, 
                                           neg_nw_10$WGT)/1000,
                   res = weighted.median(neg_nw_10$HOUSES, 
                                         neg_nw_10$WGT)/1000,
                   nonres = weighted.median(neg_nw_10$NONRES, 
                                            neg_nw_10$WGT)/1000,
                   business = weighted.median(neg_nw_10$BUS, 
                                              neg_nw_10$WGT)/1000,
                   othnfin = weighted.median(neg_nw_10$OTHNFIN, 
                                             neg_nw_10$WGT)/1000)



assets_med_07 <- data.frame(transaction = weighted.median(neg_nw_07$LIQ,
                                                          neg_nw_07$WGT)/1000,
                  securities = weighted.median(neg_nw_07$SECURITIES, 
                                               neg_nw_07$WGT)/1000,
                   ret = weighted.median(neg_nw_07$RETQLIQ, 
                                         neg_nw_07$WGT)/1000,
                   othfin = weighted.median(neg_nw_07$OTHFIN_COMPLETE,
                                            neg_nw_07$WGT)/1000,
                   vehic = weighted.median(neg_nw_07$VEHIC, 
                                           neg_nw_07$WGT)/1000,
                   res = weighted.median(neg_nw_07$HOUSES, 
                                         neg_nw_07$WGT)/1000,
                   nonres = weighted.median(neg_nw_07$NONRES, 
                                            neg_nw_07$WGT)/1000,
                   business = weighted.median(neg_nw_07$BUS, 
                                              neg_nw_07$WGT)/1000,
                   othnfin = weighted.median(neg_nw_07$OTHNFIN, 
                                             neg_nw_07$WGT)/1000)

```


## Asset holding rate by year

```{r assets holding}


assets_pct_19 <- data.frame(trans = sum(neg_nw_19$LIQ > 0)/nrow(neg_nw_19),
           sec = sum(neg_nw_19$SECURITIES > 0)/nrow(neg_nw_19),
           ret = sum(neg_nw_19$RETQLIQ > 0)/nrow(neg_nw_19),
           othfin = sum(neg_nw_19$OTHFIN_COMPLETE > 0)/nrow(neg_nw_19),
           veh = sum(neg_nw_19$VEHIC > 0)/nrow(neg_nw_19),
           res = sum(neg_nw_19$HOUSES > 0)/nrow(neg_nw_19),
           nres = sum(neg_nw_19$NONRES > 0)/nrow(neg_nw_19),
           bus = sum(neg_nw_19$BUS > 0)/nrow(neg_nw_19),
           othnfin = sum(neg_nw_19$OTHNFIN > 0)/nrow(neg_nw_19))
           

assets_pct_16 <- data.frame(trans = sum(neg_nw_16$LIQ > 0)/nrow(neg_nw_16),
           sec = sum(neg_nw_16$SECURITIES > 0)/nrow(neg_nw_16),
           ret = sum(neg_nw_16$RETQLIQ > 0)/nrow(neg_nw_16),
           othfin = sum(neg_nw_16$OTHFIN_COMPLETE > 0)/nrow(neg_nw_16),
           veh = sum(neg_nw_16$VEHIC > 0)/nrow(neg_nw_16),
           res = sum(neg_nw_16$HOUSES > 0)/nrow(neg_nw_16),
           nres = sum(neg_nw_16$NONRES > 0)/nrow(neg_nw_16),
           bus = sum(neg_nw_16$BUS > 0)/nrow(neg_nw_16),
           othnfin = sum(neg_nw_16$OTHNFIN > 0)/nrow(neg_nw_16))


assets_pct_13 <- data.frame(trans = sum(neg_nw_13$LIQ > 0)/nrow(neg_nw_13),
           sec = sum(neg_nw_13$SECURITIES > 0)/nrow(neg_nw_13),
           ret = sum(neg_nw_13$RETQLIQ > 0)/nrow(neg_nw_13),
           othfin = sum(neg_nw_13$OTHFIN_COMPLETE > 0)/nrow(neg_nw_13),
           veh = sum(neg_nw_13$VEHIC > 0)/nrow(neg_nw_13),
           res = sum(neg_nw_13$HOUSES > 0)/nrow(neg_nw_13),
           nres = sum(neg_nw_13$NONRES > 0)/nrow(neg_nw_13),
           bus = sum(neg_nw_13$BUS > 0)/nrow(neg_nw_13),
           othnfin = sum(neg_nw_13$OTHNFIN > 0)/nrow(neg_nw_13))


assets_pct_10 <- data.frame(trans = sum(neg_nw_10$LIQ > 0)/nrow(neg_nw_10),
           sec = sum(neg_nw_10$SECURITIES > 0)/nrow(neg_nw_10),
           ret = sum(neg_nw_10$RETQLIQ > 0)/nrow(neg_nw_10),
           othfin = sum(neg_nw_10$OTHFIN_COMPLETE > 0)/nrow(neg_nw_10),
           veh = sum(neg_nw_10$VEHIC > 0)/nrow(neg_nw_10),
           res = sum(neg_nw_10$HOUSES > 0)/nrow(neg_nw_10),
           nres = sum(neg_nw_10$NONRES > 0)/nrow(neg_nw_10),
           bus = sum(neg_nw_10$BUS > 0)/nrow(neg_nw_10),
           othnfin = sum(neg_nw_10$OTHNFIN > 0)/nrow(neg_nw_10))


assets_pct_07 <- data.frame(trans = sum(neg_nw_07$LIQ > 0)/nrow(neg_nw_07),
           sec = sum(neg_nw_07$SECURITIES > 0)/nrow(neg_nw_07),
           ret = sum(neg_nw_07$RETQLIQ > 0)/nrow(neg_nw_07),
           othfin = sum(neg_nw_07$OTHFIN_COMPLETE > 0)/nrow(neg_nw_07),
           veh = sum(neg_nw_07$VEHIC > 0)/nrow(neg_nw_07),
           res = sum(neg_nw_07$HOUSES > 0)/nrow(neg_nw_07),
           nres = sum(neg_nw_07$NONRES > 0)/nrow(neg_nw_07),
           bus = sum(neg_nw_07$BUS > 0)/nrow(neg_nw_07),
           othnfin = sum(neg_nw_07$OTHNFIN > 0)/nrow(neg_nw_07))

```


```{r median debt overall}

debts_med_19 <- data.frame(RESDBT = weighted.median(neg_nw_19$RESDBT,
                                                    neg_nw_19$WGT)/1000,
                  MRTHEL = weighted.median(neg_nw_19$MRTHEL, 
                                               neg_nw_19$WGT)/1000,
                   CCBAL = weighted.median(neg_nw_19$CCBAL, 
                                         neg_nw_19$WGT)/1000,
                   EDN_INST = weighted.median(neg_nw_19$EDN_INST,
                                            neg_nw_19$WGT)/1000,
                   VEH_INST = weighted.median(neg_nw_19$VEH_INST, 
                                           neg_nw_19$WGT)/1000,
                   OTHDBT = weighted.median(neg_nw_19$OTHDBT,
                                            neg_nw_19$WGT)/1000)


debts_med_16 <- data.frame(RESDBT = weighted.median(neg_nw_16$RESDBT,
                                                    neg_nw_16$WGT)/1000,
                  MRTHEL = weighted.median(neg_nw_16$MRTHEL, 
                                               neg_nw_16$WGT)/1000,
                   CCBAL = weighted.median(neg_nw_16$CCBAL, 
                                         neg_nw_16$WGT)/1000,
                   EDN_INST = weighted.median(neg_nw_16$EDN_INST,
                                            neg_nw_16$WGT)/1000,
                   VEH_INST = weighted.median(neg_nw_16$VEH_INST, 
                                           neg_nw_16$WGT)/1000,
                   OTHDBT = weighted.median(neg_nw_16$OTHDBT,
                                            neg_nw_16$WGT)/1000)


debts_med_13 <- data.frame(RESDBT = weighted.median(neg_nw_13$RESDBT,
                                                    neg_nw_13$WGT)/1000,
                  MRTHEL = weighted.median(neg_nw_13$MRTHEL, 
                                               neg_nw_13$WGT)/1000,
                   CCBAL = weighted.median(neg_nw_13$CCBAL, 
                                         neg_nw_13$WGT)/1000,
                   EDN_INST = weighted.median(neg_nw_13$EDN_INST,
                                            neg_nw_13$WGT)/1000,
                   VEH_INST = weighted.median(neg_nw_13$VEH_INST, 
                                           neg_nw_13$WGT)/1000,
                   OTHDBT = weighted.median(neg_nw_13$OTHDBT,
                                            neg_nw_13$WGT)/1000)


debts_med_10 <- data.frame(RESDBT = weighted.median(neg_nw_10$RESDBT,
                                                    neg_nw_10$WGT)/1000,
                  MRTHEL = weighted.median(neg_nw_10$MRTHEL, 
                                               neg_nw_10$WGT)/1000,
                   CCBAL = weighted.median(neg_nw_10$CCBAL, 
                                         neg_nw_10$WGT)/1000,
                   EDN_INST = weighted.median(neg_nw_10$EDN_INST,
                                            neg_nw_10$WGT)/1000,
                   VEH_INST = weighted.median(neg_nw_10$VEH_INST, 
                                           neg_nw_10$WGT)/1000,
                   OTHDBT = weighted.median(neg_nw_10$OTHDBT,
                                            neg_nw_10$WGT)/1000)



debts_med_07 <- data.frame(RESDBT = weighted.median(neg_nw_07$RESDBT,
                                                    neg_nw_07$WGT)/1000,
                  MRTHEL = weighted.median(neg_nw_07$MRTHEL, 
                                               neg_nw_07$WGT)/1000,
                   CCBAL = weighted.median(neg_nw_07$CCBAL, 
                                         neg_nw_07$WGT)/1000,
                   EDN_INST = weighted.median(neg_nw_07$EDN_INST,
                                            neg_nw_07$WGT)/1000,
                   VEH_INST = weighted.median(neg_nw_07$VEH_INST, 
                                           neg_nw_07$WGT)/1000,
                   OTHDBT = weighted.median(neg_nw_07$OTHDBT,
                                            neg_nw_07$WGT)/1000)

```


```{r debt holding}


debt_pct_19 <- data.frame(RESDBT = sum(neg_nw_19$RESDBT > 0)/nrow(neg_nw_19),
           MRTHEL = sum(neg_nw_19$MRTHEL > 0)/nrow(neg_nw_19),
           CCBAL = sum(neg_nw_19$CCBAL > 0)/nrow(neg_nw_19),
           EDN_INST = sum(neg_nw_19$EDN_INST > 0)/nrow(neg_nw_19),
           VEH_INST = sum(neg_nw_19$VEH_INST > 0)/nrow(neg_nw_19),
           OTHDBT = sum(neg_nw_19$OTHDBT > 0)/nrow(neg_nw_19))


debt_pct_16 <- data.frame(RESDBT = sum(neg_nw_16$RESDBT > 0)/nrow(neg_nw_16),
           MRTHEL = sum(neg_nw_16$MRTHEL > 0)/nrow(neg_nw_16),
           CCBAL = sum(neg_nw_16$CCBAL > 0)/nrow(neg_nw_16),
           EDN_INST = sum(neg_nw_16$EDN_INST > 0)/nrow(neg_nw_16),
           VEH_INST = sum(neg_nw_16$VEH_INST > 0)/nrow(neg_nw_16),
           OTHDBT = sum(neg_nw_16$OTHDBT > 0)/nrow(neg_nw_16))

debt_pct_13 <- data.frame(RESDBT = sum(neg_nw_13$RESDBT > 0)/nrow(neg_nw_13),
           MRTHEL = sum(neg_nw_13$MRTHEL > 0)/nrow(neg_nw_13),
           CCBAL = sum(neg_nw_13$CCBAL > 0)/nrow(neg_nw_13),
           EDN_INST = sum(neg_nw_13$EDN_INST > 0)/nrow(neg_nw_13),
           VEH_INST = sum(neg_nw_13$VEH_INST > 0)/nrow(neg_nw_13),
           OTHDBT = sum(neg_nw_13$OTHDBT > 0)/nrow(neg_nw_13))


debt_pct_10 <- data.frame(RESDBT = sum(neg_nw_13$RESDBT > 0)/nrow(neg_nw_10),
           MRTHEL = sum(neg_nw_10$MRTHEL > 0)/nrow(neg_nw_10),
           CCBAL = sum(neg_nw_10$CCBAL > 0)/nrow(neg_nw_10),
           EDN_INST = sum(neg_nw_10$EDN_INST > 0)/nrow(neg_nw_10),
           VEH_INST = sum(neg_nw_10$VEH_INST > 0)/nrow(neg_nw_10),
           OTHDBT = sum(neg_nw_10$OTHDBT > 0)/nrow(neg_nw_10))


debt_pct_07 <- data.frame(RESDBT = sum(neg_nw_13$RESDBT > 0)/nrow(neg_nw_07),
           MRTHEL = sum(neg_nw_07$MRTHEL > 0)/nrow(neg_nw_07),
           CCBAL = sum(neg_nw_07$CCBAL > 0)/nrow(neg_nw_07),
           EDN_INST = sum(neg_nw_07$EDN_INST > 0)/nrow(neg_nw_07),
           VEH_INST = sum(neg_nw_07$VEH_INST > 0)/nrow(neg_nw_07),
           OTHDBT = sum(neg_nw_07$OTHDBT > 0)/nrow(neg_nw_07))

debts_time <- data.frame(year = c("2007", "2010", "2013", "2016", "2019"),
                         rbind(debt_pct_07, debt_pct_10, debt_pct_13, 
                               debt_pct_16, debt_pct_19))


write.csv(debts_time,  paste(getwd(), '\\csv-exports', 
                '\\debts_time.csv', 
                sep=''))

```


## Education level

I calculate the share of households by education level for those with negative net worth and overall. See here for the level of education the numbered categories correspond to: https://sda.berkeley.edu/sdaweb/docs/scfcomb2019/DOC/hcbk0001.htm#EDCL. 

```{r Education}

# Education levels for all HHs and net worth HHs

wpct(scf_19$EDCL, scf_19$WGT)
wpct(neg_nw_19$EDCL, neg_nw_19$WGT)

```


## Income to Debt Ratio

```{r Income to debt ratio}

# Debt to income ratio compared to overall population

weighted.median(neg_nw_19$DEBT2INC, neg_nw_19$WGT)
weighted.median(scf_19$DEBT2INC, scf_19$WGT)

```

## Deliquent payments

```{r Late payments}


# Late payment in the last year 

wpct(scf_19$LATE, scf_19$WGT)
wpct(neg_nw_19$LATE, neg_nw_19$WGT)


# Payment past 60 days due 

wpct(scf_19$LATE60, scf_19$WGT)
wpct(neg_nw_19$LATE60, neg_nw_19$WGT)

```


